On centrality of orthogonal $\Kt$

### 2016-04-25 paper.tex
Proofread one more page

In \cite{Sus} Suslin showed for $n\geq 3$ that the elementary group $\E(n, R)$ coincides with the subgroup of $\GL(n,\,R)$ generated by matrices
of the form $t(u,\,v)=1+uv^t$ where $u\in\Um(n,\,R)$, $v\in R^n$ and $u^tv=0$. Here $u^t$ stands for the transpose of $u$.
This result clearly implies that $\E(n,\,R)$ is normal inside $\GL(n,\,R)$.
Afterwards, W.~van~der~Kallen developed Suslin's ideas and showed for $n\geq4$ that the Steinberg group $\St(n,\,R)$ is isomorphic to the group presented by generators
$$\{X(u,\,v)\mid u\in\Um(n,\,R),\ v\in R^n,\ u^tv=0\}$$ and the following list of relations (see~\cite{vdK}):
\setcounter{equation}{0} \renewcommand{\theequation}{K\arabic{equation}}
\begin{align}
&X(u,\,v)X(u,\,w)=X(u,\,v+w), \label{add1} \\
&X(u,\,v)X(u',\,v')X(u,\,v)\inv=X(t(u,\,v)u',\,t(v,\,u)\inv v'). \label{conj1}
\end{align}
This presentation clearly implies that $\phi\colon\St(n,\,R)\rightarrow\E(n,\,R)$ is a central extension.
Notice that our notation for the generators slightly differs from that of~\cite{vdK}, e.\,g. our elements $X(u, v)$ correspond to van der Kallen's $(u, v^t)$.

@@

Next, we show that for a splitting ideal $I\trianglelefteq R$ the group $\St^*(n,\,R,\,I)$ is isomorphic to $\St(n,\,R,\,I)=\Ker\big(\St(n,\,R)\epi\St(n,\,R/I)\big)$. With this end we construct two inverse homomorphisms
$$
\xymatrix{
\St^*(n,\,R,\,I)\ar@<0.5ex>[r]^{\iota}&\St(n,\,R,\,I)\ar@<0.5ex>[l]^{\kappa}.
}
$$

Firstly, we define $\iota$ as a homomorhism to an absolute group
$$
\xymatrix{
\St^*(n,\,R,\,I)\ar@<0.0ex>[r]^{\iota}&\St(n,\,R).
}
$$
We do not need an assumption that $I$ is a splitting ideal for that.

Clearly, we map elements $F(u,\,v)$ to van der Kallen's another generators $X(u,\,v)$. To find images for $S(u,\,v)$ as well we recall that van der Kallen in~\cite[3.8--3.10]{vdK} defines also elements
$$
x(u,\,v)\in\St(n,\,R),\ u^tv=0,\ u_i=0\,\text{ or }\,v_i=0\,\text{ for some }\,1\leq i\leq n.
$$
Now, ``transpose'' the definition~\cite[3.13]{vdK}.

\begin{df}
For $u\in R^n$, $v\in\E(n,\,R)e_1$, $u^tv=0$, consider the set $\overline Y(u,\,v)\subseteq\St(n,\,R)$ consisiting of all elements $y\in\St(n,\,R)$ which could be decomposed as a product $\prod x(w^k,\,v)$, where $\sum w^k=u$ and $w^k=(e_pv_q-e_qv_p)c^k$ for some $c^k\in R$, $1\leq p\neq q\leq n$.
\end{df}

Since coloumns of elementary matrices are unimodular, $\overline Y(u,\,v)$ is not empty (see~\cite[3.1--3.2]{vdK}). Obviously, for $x\in\overline Y(u,\,v)$ and $y\in\overline Y(w,\,v)$ holds $xy\in\overline Y(u+w,\,v)$.

Repeating~\cite[3.14--3.15]{vdK} verbatim one shows that

\begin{lm}
for $g\in\St(n,\,R)$ holds
\begin{enumerate}
\item
$g\overline Y(u,\,v)g\inv\subseteq\overline Y(\phi(g)u,\,\phi(g)^*v)$;
\item
$\overline Y(u,\,v)$ consists of exactly one element.
\end{enumerate}
\end{lm}

The only element of $\overline Y(u,\,v)$ we denote by $Y(u,\,v)$. We map elements $S(u,\,v)\in\St^*(n,\,R,\,I)$ to $Y(u,\,v)$. To show that such a homomorphism is well-defined we check that $X(u,\,v)$ and $Y(u,\,v)$ satisfy relations R1--R4, but R1--R3 are already checked.

\begin{lm}
Consider $(u,\,v^t)=(Me_1,\,e_2^tM\inv)$, $M\in\E(n,\,R)$, $a\in I$. Then $X(u,\,va)=Y(ua,\,v)$.
\end{lm}

\begin{proof}
It suffices to show that $X(e_1,\,e_2a)=Y(e_1a,\,e_2)$. Compute the commutator $[Y(-e_3,\,e_2),\,X(e_1,\,e_3a)]$ in two ways:
$$
Y(-e_3,\,e_2)X(e_1,\,e_3a)Y(-e_3,\,e_2)\inv\cdot X(e_1,\,-e_3a)=X(e_1,\,e_2a),
$$
and
$$
Y(-e_3,\,e_2)\cdot\,X(e_1,\,e_3a)Y(e_3,\,e_2)X(e_1,\,e_3a)\inv=Y(e_1a,\,e_2).
$$
\end{proof}

Therefore, the homomorphism $\iota\colon\St^*(n,\,R,\,I)\rightarrow\St(n,\,R)$ sending $F(u,\,v)$ to $X(u,\,v)$ and $S(u,\,v)$ to $Y(u,\,v)$ is well-defined. Denote $\rho_I\colon R\epi R/I$, then one can easily see that induced morphism on Steinberg groups 
$$\rho_I^*\colon\St(n,\,R)\rightarrow\St(n,\,R/I)$$
 sends $X(u,\,v)$ to $X(\rho_I(u),\,\rho_I(v))$ and $Y(u,\,v)$ to $Y(\rho_I(u),\,\rho_I(v))$. Observe that $\E(n,\,R)\epi\E(n,\,R/I)$ is surjective so that for $x\in\E(n,\,R)e_1$ holds that $\rho_I(x)\in\E(n,\,R/I)e_1$. It follows from additivity relations that $X(u,\,0)=1=Y(0,\,v)$, thus $\mathrm{Im}(\iota)\subseteq\Ker(\rho_I^*)=\St(n,\,R,\,I)$.

>>>>

In \cite{Sus} Suslin showed for $n\geq 3$ that the elementary group $\E(n, R)$ coincides with the subgroup of $\GL(n,\,R)$ generated by matrices
of the form $t(u,\,v)=1+uv^t$ where $u\in\Um(n,\,R)$, $v\in R^n$ and $u^tv=0$. Here $u^t$ stands for the transpose of $u$.
This result clearly implies that $\E(n,\,R)$ is normal inside $\GL(n,\,R)$.
Afterwards, W.~van~der~Kallen developed Suslin's ideas and showed for $n\geq4$ that the Steinberg group $\St(n,\,R)$ is isomorphic to the group presented by generators
$$\{X(u,\,v)\mid u\in\Um(n,\,R),\ v\in R^n,\ u^tv=0\}$$ and the following list of relations (see~\cite[Theorem~1]{vdK}):
\setcounter{equation}{0} \renewcommand{\theequation}{K\arabic{equation}}
\begin{align}
&X(u,\,v)X(u,\,w)=X(u,\,v+w), \label{add1} \\
&X(u,\,v)X(u',\,v')X(u,\,v)\inv=X(t(u,\,v)u',\,t(v,\,u)\inv v'). \label{conj1}
\end{align}
This presentation clearly implies that $\phi\colon\St(n,\,R)\rightarrow\E(n,\,R)$ is a central extension.
Notice that our notation for the generators of $\St(n, R)$ slightly differs from that of~\cite{vdK}, e.\,g. we parametrize elements $X(u, v)$ with two columns
rather than with one column and one row. Our generator $X(u,v)$ corresponds to van der Kallen's generator $(u, v^t)$.

@@

Our next goal is to show for a splitting ideal $I\trianglelefteq R$ that the group $\St^*(n,\,R,\,I)$ is isomorphic to $\St(n,\,R,\,I)=\Ker\big(\St(n,\,R)\epi\St(n,\,R/I)\big)$.
With this end we construct two mutually inverse homomorphisms
$$\xymatrix{\St^*(n,\,R,\,I)\ar@<0.5ex>[r]^{\iota}&\St(n,\,R,\,I)\ar@<0.5ex>[l]^{\kappa}.}$$

Firstly, we define $\iota$ as a homomorphism to the absolute group (for which we de not need the assumption that $I$ is splitting).
$$\xymatrix{\St^*(n,\,R,\,I)\ar@<0.0ex>[r]^{\iota}&\St(n,\,R).}$$

Clearly, we must map the elements $F(u,\,v)$ to van der Kallen's generators $X(u,\,v)$.
Now we should find the images for the elements $S(u,\,v)$ as well.
First of all, recall that van der Kallen in~\cite[3.8--3.10]{vdK} constructs the following elements
$$x(u,\,v)\in\St(n,\,R),\ u^tv=0,\ u_i=0\,\text{ or }\,v_i=0\,\text{ for some }\,1\leq i\leq n.$$
The following definition is the ``transpose'' of \cite[3.13]{vdK}.
\begin{df} For $u\in R^n$, $v\in\E(n,\,R)e_1$, $u^tv=0$, consider the set $\overline Y(u,\,v)\subseteq\St(n,\,R)$
 consisiting of all elements $y\in\St(n,\,R)$ that admit the decomposition $y=\prod x(w^k,\,v)$, 
 where $\sum u^k=u$ and $u^k=c^k \cdot (e_pv_q-e_qv_p)$ for some $c^k\in R$, $1\leq p\neq q\leq n$. \end{df}

Since columns of elementary matrices are unimodular, one can find $w\in R^t$ such that $w^tv=1$.
It is not hard to check that $u = (w^t\cdot v)\cdot u = \sum_{p<q}u_{pq}$, where $u_{pq} = (u_pw_q - u_qw_p) \cdot (v_qe_p - v_pe_q)\in{}\!R^n$,
therefore $\overline Y(u,\,v)$ is not empty (cf.~\cite[3.1--3.2]{vdK}).
Obviously, for $x\in\overline Y(u,\,v)$ and $y\in\overline Y(w,\,v)$ one has $xy\in\overline Y(u+w,\,v)$.

Repeating~\cite[3.14--3.15]{vdK} verbatim one shows the following.
\begin{lm} For $g\in\St(n,\,R)$ one has
\begin{enumerate} \item $g\overline Y(u,\,v)g\inv\subseteq\overline Y(\phi(g)u,\,\phi(g)^*v)$;
                  \item $\overline Y(u,\,v)$ consists of exactly one element. \end{enumerate} \end{lm}

We denote the only element of $\overline Y(u,\,v)$ by $Y(u,\,v)$. 
Now we can define $\iota$ by sending $S(u,\,v)\in\St^*(n,\,R,\,I)$ to $Y(u,\,v)$. 
To show that the map $\iota$ is well-defined we should check that the elements $X(u,\,v)$ and $Y(u,\,v)$ satisfy the relations 
\eqref{add4}--\eqref{coinc} (with $F$'s and $S$'s replaced with $X$'s and $Y$'s).
Notice that the validity of relations \eqref{add4}--\eqref{conj3} has already been checked.
\begin{lm} For $(u,\,v^t)=(Me_1,\,e_2^tM\inv)$, $M\in\E(n,\,R)$, $a\in I$ one has $X(u,\,va)=Y(ua,\,v)$. \end{lm}
\begin{proof}
In view of the above lemma we only need to show that $X(e_1,\,e_2a)=Y(e_1a,\,e_2)$.
This equality can be obtained by computing the commutator $[Y(-e_3,\,e_2),\,X(e_1,\,e_3a)]$ in two ways:
$$ Y(-e_3,\,e_2)X(e_1,\,e_3a)Y(-e_3,\,e_2)\inv\cdot X(e_1,\,-e_3a)=X(e_1,\,e_2a), \text{ and} $$
$$ Y(-e_3,\,e_2)\cdot\,X(e_1,\,e_3a)Y(e_3,\,e_2)X(e_1,\,e_3a)\inv=Y(e_1a,\,e_2). \qedhere $$
\end{proof}

We can sum up the above discussion into the following result.
\begin{lm} \label{lm:map-iota}
 For a splitting ideal $I\trianglelefteq R$ there exists a well-defined homomorphism 
 $$\iota\colon\St^*(n,\,R,\,I)\rightarrow\St(n,\,R, I) \subseteq \St(n, R)$$ 
 satisfying the equalities $\iota (F(u,\,v)) = X(u,\,v)$ and $\iota (S(u,\,v)) = Y(u,\,v)$. \end{lm}
\begin{proof}
We already know that $\iota$ is well-defined. It suffices to check that the image of $\iota$ is contained in $\St(n,\,R,I)$.
Denote the canonical projection $R\epi R/I$ by $\rho$.
Clearly, the induced group morphism $\rho^*\colon\St(n,\,R)\rightarrow\St(n,\,R/I)$ sends $X(u,\,v)$ to $X(\rho(u),\,\rho(v))$ and $Y(u,\,v)$ to $Y(\rho(u),\,\rho(v))$. 
Observe that $\E(n,\,R)\epi\E(n,\,R/I)$ is surjective, therefore for $x\in\E(n,\,R)e_1$ one has $\rho(x)\in\E(n,\,R/I)e_1$. 
From the additivity relations we get that $X(u,\,0)=1=Y(0,\,v)$, hence $\mathrm{Im}(\iota)\subseteq\Ker(\rho^*)=\St(n,\,R,\,I)$, as required. \end{proof}
